---
title: "###Storms and Severe weather"
output: html_document
---

###### Storm and other severe weather events' impact problems on humans and economic impact
###### Synopsis

In this report, I collect the data related to storm and other severe weather events across the United States, run analysis and at the conclusion of this research
I will present you with my findings.
```{r, cache=TRUE}
setwd("M:/ReproducibleResearch")
if (!(file.exists("./StormData.csv.bz2")))
  {
     url <- "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
     download.file(url, destfile = "./StormData.csv.bz2")
  } else storm.data <- read.csv(bzfile("StormData.csv.bz2"), header = TRUE, sep = ",")

if (!"storm.data" %in% ls())
  { 
     storm.data <- read.csv(bzfile("StormData.csv.bz2"), header = TRUE, sep = ",")
  }
head(storm.data)
str(storm.data) # looking at the dataframe structure 
```

###### Subsetting the data to columns of interests (fatalities, injuries, property  damage, and crop damage) 
```{r data subset}
library(dplyr)
storm.subset <- select(storm.data, BGN_DATE, FATALITIES, INJURIES, EVTYPE, PROPDMG, CROPDMG, STATE)
```
###### Events resulting in fatalities
```{r fatalities data}
storm.fatalities <- aggregate(storm.subset$FATALITIES, by = list(storm.subset$EVTYPE), FUN = sum, na.rm = TRUE)
names(storm.fatalities) <- c("EVENT_TYPE", "FATALITIES") # renaming the columns
# Sorting data from the highest to lowest
# Total fatalities by events
fatalities.sorted <- arrange(storm.fatalities, desc(FATALITIES))
# Top 10 death-causing events
head(fatalities.sorted, 10) 
# Bottom 10 death-causing events
tail(fatalities.sorted, 10)
```
##### Graphical representation of the number of deaths by events
```{r death graph}
lbls <- paste(fatalities.sorted$EVENT_TYPE, "\n", fatalities.sorted$FATALITIES, sep = "")
pie(fatalities.sorted$FATALITIES, labels = lbls, main = "Fatalities by events across the US")
```